The robot who pretend-played with me

### 2017-04-06 paper.tex
Analysis of the freeplay sandbox

The study of the emergence of social behaviour is often intrinsically limited by
the nature of the experimental tasks. We introduce a new
experimental paradigm based on free play interactions. Where traditional
socio-cognitive tasks are either toy scenarios (\ie they do not adequately
mirror real-world situations) or simple, constrained tasks that do not reflect
the complexity and dynamics of real-world interactions, free play social
interactions exhibit a rich set of natural cognitive and social dynamics that
are desirable for this research.

@@

To mitigate the technical challenges presented by free play interactions (like
high dimensionality and uncertainties), we have developped a  \emph{free
play sandbox}~\ref{fig|setup}: open-ended and non-directive activities, yet sufficiently
defined to be reproducible; focus on abstract socio-cognitive facets (robot
perception and manipulation are simplified); well suited for qualitative and
quantitative analysis using metrics like Słowinski’s Individual Motor Signature
(for behavioural alignment), Ballard’s coding of children’s free-play
interactions, with-me-ness (for assessment of co-engagment).

\subsection{Free play task}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{sandbox-release-cheetah}
    \caption{Possible play situation: the child decides to release the cheetah
    and creates a story where the cheetah scares away the visitors.}
    \label{}
\end{figure}

>>>>

The study of the emergence of social behaviour is often intrinsically limited by
the nature of the experimental tasks. Where traditional
socio-cognitive tasks are either toy scenarios (\ie they do not adequately
mirror real-world situations) or simple, constrained tasks that do not reflect
the complexity and dynamics of real-world interactions, free play social
interactions exhibit a rich set of natural cognitive and social dynamics that
are desirable for this research. We introduce a new
experimental paradigm based on free play interactions.

@@

To mitigate the technical challenges presented by free play interactions in
their most unconstrained version (like
high dimensionality and uncertainties), we have developped a \emph{free play
sandbox}, Figure~\ref{fig|setup}: a large touchscreen (60cm $\times$ 33cm) is
used as an interactive surface (\emph{sandtray}
paradigm~\cite{baxter2012touchscreen}). Two agents (either a pair of children or
one child and one robot) play together by freely moving on a 2D plane items
(characters and blocks, see Figure~\ref{fig|sandbox}). A background image
depicts a generic empty environment, with different symbolic zones (water,
grass, path, bushes...). The players do not have any particular task to
complete, they are simply invited to freely play.

children can engage in open-ended and
non-directive play situations, yet sufficiently
defined to be reproducible; focus on abstract socio-cognitive facets (robot
perception and manipulation are simplified); well suited for qualitative and
quantitative analysis using metrics like Słowinski’s Individual Motor Signature
(for behavioural alignment), Ballard’s coding of children’s free-play
interactions, with-me-ness (for assessment of co-engagment).

\subsection{Analysis of the free play sandbox}

Our free play sandbox elicits a loosely structured play situation. The
situation is indeed essentially aimless (free play). The goals and play episodes
that we observe during the interactions are not pre-defined, and essentially
unknown until they are decided by the players. In this sense, the free play
sandbox permits and elicits what is fundamentally \emph{unstructured play}.

However, a second look reveals some important structuring elements.

First, the physical bounds of the sandbox (an interactive table) limit the
play zone to a well defined and relatively small area. As a consequence,
children are mostly static (they are sitting in front of the table) and their
primary form of interaction is 2D pick and place of items by drag and dropping
them on the screen.

Second, the game items themselves structure the game scenarios. Items are either
iconic characters (animals or children) or plain blocks
(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
(like 'crocodiles like water and eat children'), while the later have little
semantics associated. The game background, with its recognizable zones, also
elicit a particular type of games (like building a zoo or pretending we explore
the savannah).

Finally, the social setting (two players facing each other) implicitely invites
social interaction: onlooker play, for instance, is unlikely to be observed as
the children are physically engaged with the game and the other participant
(they are sitting in front of game, close to each other). The three higher forms
of social play listed by Parten (parallel play, associative play and cooperative
play) are however possible with this setup.

Two more elements (the role and place of the experimenter, and the initial
prompt given to the participant) have an important impact on the shape of the
play situation. We discuss separately these two points in the description of the
experimental procedure.

Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
actual play situations are not known and might change several times during the
interaction; the game actions, even though all based on a single modality (picking and
placing game items), are unlimited; the social interactions between participants
are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
unconstrainted. However, the broad domain of the play situations and the range of
possible social interactions is bound by the physical bounds of the play zone
and the theme of the game items.

\subsection{Forms of play elicited by the sandbox}

\begin{itemize}
    \item storytelling
    \item creative play
    \item pretend play
\end{itemize}

\paragraph{From unstructured to structured play}

@@

\paragraph{Pretence}

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{sandbox-release-cheetah}
    \caption{Possible play situation: the children pretend to release the cheetah
    and create a story where the cheetah scares away the visitors.}
    \label{}
\end{figure}
